### 2020-02-20 Common/chapters/paper_reproduction.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest dataset available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.

@@

\section{Process overview}
Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches.
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved. Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC. As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
This section describes all processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (section \ref{sec:classification_challenge}).

@@

The PROSTATEx dataset comes with two CSV files for the training set. The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can be associated with the same patient (ProxID). The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding. Important labels are "ProxID" (patient ID), "fid" (finding ID $\in [1,\infty]$), "ClinSig" (clinical significance, TRUE or FALSE), "DCMSerNumber" (digit before the dash in the folder name containing DICOM files), "ijk" (position of the lesion: slice number $k$ at coordinates $(i,j)$, $i,j,k \in [0,\infty]$) and "VoxelSpacing" (3-dimensional vector representing the correspondance between a pixel and the space it occupies in the real world). Both CSV files are complementary to each other. Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}. The only difference is the absence of clinical significance. Finally, images are in the DICOM file format (see section \ref{sec:DICOM}).

\subsubsection{Methodology}
The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors didn't mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60x$) in order to create balanced training and validation sets.

@@

\begin{algorithm}
    \caption{PROSTATEx preprocessing}
    \label{alg:PROSTATEx_preprocessing}
    \begin{algorithmic}[1] % The number tells where the line numbering should start
        \Procedure{main}{$dataset\_folder, findings\_CSV, slices\_CSV, output\_folder$}
        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
        		\State $findings \gets read\_CSV(findings\_CSV)$ \Comment{ProstateX-Findings-Train.csv}
        		\State $slices \gets read\_CSV(slices\_CSV)$ \Comment{ProstateX-Images-Train.csv}
            \State $meta \gets merge(findings, slices)$\Comment{Both CSV files are complementary to each other.}\\
            \For{$row$ in $meta$}
            		\State $patient\_id \gets row["ProxID"]$
            		\State $finding\_id \gets row["fid"]$
            		\State $mri\_type\_number \gets row["DCMSerNumber"]$
            		\State $clinical\_significance \gets row["ClinSig"]$
            		\State $img\_i, img\_j, img\_k \gets row["ijk"]$
            		\State $slice\_number \gets img\_k + 1$ \Comment{CSV indexing in $[0,\infty]$, DICOM in $[1,\infty]$}\\
                \For{$visit\ in\ patient\_id$'s folder}
                		\For{$mri\_type$ in $visit$}
                			\If {$mri\_type$ starts with "$mri\_type\_number$-"}
                				\For{$dicom\_file$ in $mri\_type$}
                					\If {$slice\_number == dicom\_file.InstanceNumber$}
                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
                						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
                					\EndIf
                				\EndFor
                			\EndIf
            			\EndFor
            		\EndFor
            \EndFor
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection{From NumPy arrays to augmented stacked images}
\label{sec:numpyToAugmentedStacked}
This step can be split into two substeps: aligning and stacking images before augmenting them. Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution. In fact, T2-weighted images have a much higher resolution than the DWI and ADC images (in this dataset at least). Without this operation, a single pixel on a DWI image would have covered a lot more tissue than a pixel of the corresponding T2-image. To perform the alignment, the lesion was localized on the three sequences than to the CSV file. Then, the goal was to crop a large patch which contained the same amount of tissue on the three images. This patch had to be centered on the lesion to ease the augmentation process. So, a fixed patch size was defined for the T2-image since its resolution was the highest. Then, using the voxel spacing information, the patch size required to cover the same amount of tissue was computed for the two other sequences (DWI, ADC). Finally, the three images were cropped, resized to the same resolution and stacked. Stacking images consists in putting each single image (grayscale) into a single array, as three different channels. At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel. This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible on the three images.

@@

\noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient. According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately help the final diagnosis. In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array. Then, the mean and standard devation of the array were computed. These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient. Same for this patient's DWI and ADC images. The same process was applied to all patients separately.

Finally, the dataset was augmented in the same way as the authors. Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel). These techniques alone allowed to create a sufficient amount of data. Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not made use of. Images were exported as NumPy arrays.

\subsubsection{From NumPy arrays to augmented non-stacked images}
Another way of processing and using the data was experimented with. Instead of stacking the images, each image was cropped and augmented separately. At the end, each image was given as separate input to the neural network.

@@

\subsubsection{Experimental setup}
Keeping the exact same hyperparameters as the original paper lead the experiment to disappointing results. Consequently, new hyperparameters had to be found. A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen. In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set. At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2. This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\

\subsection{Training verification}
In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. In this case, large gradients are computed, which leads to huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).

@@

\subsection{Results}
\label{ref:sec:results}
Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$.\\
Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71. This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.

@@

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% as it was done in the previous section. First, the changes made to reach this goal are detailed, then the raw results are presented and finally the latter are interpreted.
\subsection{Training the neural network with the whole dataset}
\subsection{Results on the challenge test set}

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge_full_dataset.png}
\caption{Metrics on the training using all data available as training set}
\label{fig:challenge_all_results}
\end{figure}

>>>>

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest dataset available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.

@@

\section{Process overview}
Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches.
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved. Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC. As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
This section describes all processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).

@@

The PROSTATEx dataset comes with two CSV files for the training set. The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can be associated with the same patient (ProxID). The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding. Important labels are "ProxID" (patient ID), "fid" (finding ID $\in [1,\infty]$), "ClinSig" (clinical significance, TRUE or FALSE), "DCMSerNumber" (digit before the dash in the folder name containing DICOM files), "ijk" (position of the lesion: slice number $k$ at coordinates $(i,j)$, $i,j,k \in [0,\infty]$) and "VoxelSpacing" (3-dimensional vector representing the correspondance between a pixel and the space it occupies in the real world). Both CSV files are complementary to each other. Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}. The only difference is the absence of clinical significance. Finally, images are in the DICOM file format (see Section \ref{sec:DICOM}).

\subsubsection{Methodology}
The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors didn't mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.

@@

\begin{algorithm}
    \caption{PROSTATEx preprocessing}
    \label{alg:PROSTATEx_preprocessing}
    \begin{algorithmic}[1] % The number tells where the line numbering should start
        \Procedure{main}{$dataset\_folder, findings\_CSV, slices\_CSV, output\_folder$}
        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
        		\State $findings \gets read\_CSV(findings\_CSV)$ \Comment{ProstateX-Findings-Train.csv}
        		\State $slices \gets read\_CSV(slices\_CSV)$ \Comment{ProstateX-Images-Train.csv}
            \State $meta \gets merge(findings, slices)$\Comment{Both CSV files are complementary to each other.}\\
            \For{$row$ in $meta$}
            		\State $patient\_id \gets row["ProxID"]$
            		\State $finding\_id \gets row["fid"]$
            		\State $mri\_type\_number \gets row["DCMSerNumber"]$
            		\State $clinical\_significance \gets row["ClinSig"]$
            		\State $img\_i, img\_j, img\_k \gets row["ijk"]$
            		\State $slice\_number \gets img\_k + 1$ \Comment{CSV indexing in $[0,\infty]$, DICOM in $[1,\infty]$}\\
                \For{$visit\ in\ patient\_id$'s folder}
                		\For{$mri\_type$ in $visit$}
                			\If {$mri\_type$ starts with "$mri\_type\_number$-"}
                				\For{$dicom\_file$ in $mri\_type$}
                					\If {$slice\_number == dicom\_file.InstanceNumber$}
                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see Section \ref{sec:dicom_data_manipulation}}
                						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
                					\EndIf
                				\EndFor
                			\EndIf
            			\EndFor
            		\EndFor
            \EndFor
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection{From NumPy arrays to augmented stacked images}
\label{sec:numpyToAugmentedStacked}
This step can be split into two substeps: aligning and stacking images before augmenting them. Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution. In fact, T2-weighted images have a much higher resolution than the DWI and ADC images (in this dataset at least). Without this operation, a single pixel on a DWI image would have covered a lot more tissue than a pixel of the corresponding T2-image. To perform the alignment, the lesion was localized on the three sequences than to the CSV file. Then, the goal was to crop a large patch which contained the same amount of tissue on the three images. This patch had to be centered on the lesion to ease the augmentation process. So, a fixed patch size was defined for the T2-image since its resolution was the highest. Then, using the voxel spacing information, the patch size required to cover the same amount of tissue was computed for the two other sequences (DWI, ADC). Finally, the three images were cropped, resized to the same resolution and stacked. Stacking images consists in putting each single image (grayscale) into a single array, as three different channels. At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel. This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.

@@

\noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient. According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately helps the final diagnosis. In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array. Then, the mean and standard devation of the array were computed. These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient. Same for the patient's DWI and ADC images. The same process was applied to all patients separately.

Finally, the dataset was augmented in the same way as the authors. Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel). These techniques alone allowed to create a sufficient amount of data. Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used. Images were exported as NumPy arrays.

\subsubsection{From NumPy arrays to augmented non-stacked images}
Another way of processing and using the data was experimented. Instead of stacking the images, each image was cropped and augmented separately. At the end, each image was given as separate input to the neural network.

@@

\subsubsection{Experimental setup}
Keeping the exact same hyperparameters as the original paper lead the experiment to disappointing results. Consequently, new hyperparameters had to be found. A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen. In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set. At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2. Note that the model was initialized with the best one among 200 models coming from a roulette maximizing the f1-score. This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\

\subsection{Training verification}
In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. In this case, large gradients are computed, which leads to huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This implementation is based on the code of RoshanRane \cite{47}.  This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).

@@

\subsection{Results}
\label{ref:sec:results}
Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}).\\
Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reprodution_results_challenge_1}). This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.

@@

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section. First, the changes needed to reach this new goal are detailed, then the raw results are presented and finally the latter are interpreted.
\subsection{Training the neural network with the whole dataset}
The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters. It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state of the art deep learning model for prostate lesions classification.\\
As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data. In a general way, the more data are available the better will be the model. Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from previous section. Apart from this new split, the data processing and the hyperparameters keep the same. The only difference in the training is about the learning rate decay. Indeed, since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting. However, this does not cause any problem because, thanks to the results of the previous section, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied. Similarly, it is possible to conclude that in this case the best model should also be the one around epoch 17. In order to ensure this, the models at epoch 15, 17, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set. For each of these models, a csv file containing the predictions for each lesion of each patient was automatically generated and sent to the challenge.
\subsection{Results on the challenge test set}
Figure \ref{fig:challenge_all_results_experiment} presents the metrics achieved by the model trained with all available data as training set at each epoch.\\
Figure \ref{fig:results_for_all_models_challenge} shows the performance of the model saved at different epoch on the official challenge test set. The best AUC achieved on the challenge is 0.76. Figure \ref{fig:paper_reprodution_results_challenge2} illustrates the participation to

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge_full_dataset.png}
\caption{Metrics on the training using all data available as training set}
\label{fig:challenge_all_results_experiment}
\end{figure}

\begin{table}[]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
\hline
\rowcolor[HTML]{DAE8FC} 
\textbf{Epoch} & \textbf{15} & \textbf{17} & \textbf{19} & \textbf{20} & \textbf{21} & \textbf{22} & \textbf{23} & \textbf{30} \\ \hline
AUC            & 0.75        & 0.76        & 0.76        & 0.76        & 0.76        & 0.76        & 0.75        & 0.75        \\ \hline
\end{tabular}%
}
\caption{AUC of the model saved at different epochs on the official SPIE-AAPM-NCI Prostate MR classification test set }
\label{fig:results_for_all_models_challenge}
\end{table}

### 2020-02-20 Common/chapters/transfer_learning.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

\chapter{Improving performance using transfer learning}
\label{ch:transfer_learning}
%What we do: improve cancer detection/classification on medical imaging
%How we do it: transfer learning
This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data. To achieve this, transfer learning is used.\\
First, a general definition of transfer learning is given. Its utility in this specific application is discussed as it is useful but not in any case. Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied. Then, a section focuses on the processing of new datasets that are used in the final experiment. Finally, the latter and its results are discussed.

\section{Goal}
Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B. This technique makes sense when the quantity of data available is not sufficient and when the data of both datasets share common features. Both conditions are fulfilled in the case of cancer detection. In fact, medical data are highly confidential and cannot be shared easily. Provided that medical institutions agree to share it, it has to undergo an anonymization process which follows very strict rules. Then, to be usable as part of a deep learning classication task, data must be organized and described precisely. In other words, additional information such as the clinical significance, the position of the lesion, etc. must be collected and put together. This whole process requires considerable effort, which is one of the reasons why medical data is hard to find. In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.

\section{Process overview}
\label{sec:process_overview}
The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
First of all, similar datasets are collected and processed in the same way. These datasets need to share common features. For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.). On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.

>>>>

\chapter{Improving performance using transfer learning}
\label{ch:transfer_learning}
This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data. To achieve this, transfer learning is used.\\
First, a general definition of transfer learning is given. Its utility in this specific application is discussed as it is useful but not in any case. Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied. Then, a section focuses on the processing of new datasets that are used in the final experiment. Finally, the latter and its results are discussed.

\section{Goal}
Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B. This technique makes sense when the quantity of data available is not sufficient and when the data of both datasets share common features. Both conditions are fulfilled in the case of cancer detection. In fact, medical data are highly confidential and cannot be shared easily. Provided that medical institutions agree to share it, it has to undergo an anonymization process which follows very strict rules. Then, to be usable as part of a deep learning classification task, data must be organized and described precisely. In other words, additional information such as the clinical significance, the position of the lesion, etc. must be collected and put together. This whole process requires considerable effort, which is one of the reasons why medical data is hard to find. In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.

\section{Process overview}
\label{sec:process_overview}
The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
First of all, similar datasets are collected and processed in the same way. These datasets need to share common features. For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.). 
%On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
